---
title: "MPP_transplantation_expt"
author: "Jason Cosgrove"
date: '2022-07-20'
output: html_document
---



# Analysis of clone size, diversity in fate of barcoded MPPs

In this experiment we sorted MPPs from WT mice (mice were sent from IMDEA madrid) and barcoded them with the LG2.2 library. Cells were transplanted into irradiated recipients and 3 weeks later the bone marrow was harvested, barcoded cells were sorted and lysed for barcode library sequencing. 



### Step 1: Set up the workspace
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/jasoncosgrove/Dropbox (Team_Perie)/Jason/Experiments/Wet_lab/JCW24_G6PD_barcoding/LentiviralBarcoding_Analysis/inputs/")

#clear the workspace and load in the necessary packages
rm(list = ls())
library(cowplot)
library(plotrix)
library(ggplot2)
library(dplyr)


setwd("/Users/jasoncosgrove/Dropbox (Team_Perie)/Jason/Experiments/Wet_lab/JCW24_G6PD_barcoding/LentiviralBarcoding_Analysis/inputs/")

#load int he cell couunts that we will use later on for normalisation
cell_counts <- read.csv('/Users/jasoncosgrove/Dropbox (Team_Perie)/Jason/Experiments/Wet_lab/JCW24_G6PD_barcoding/LentiviralBarcoding_Analysis/cell_counts.csv')


```


```{r helper_methods, echo=FALSE}

columnNormalisation <- function(data,barcodes){
 
   data.normed <- apply(data,2, function(x) (x/sum(x))) # here we normalise for each sample
   rownames(data.normed) <- barcodes
   return(data.normed)

}

removeUninformativeClones <- function(mat){
      #remove any rows where both pos and neg are zero
    m.filtered <- mat[rowSums(mat)> 0,]
    return(m.filtered)
}

getMouseMatrix <- function(mat, mouse.id){
    
   mouse.columns <- colnames(mat)[grepl(mouse.id,colnames(mat))]
   mouse <- data.frame(mat[,c(mouse.columns)])
   m <- removeUninformativeClones(mouse)
   colnames(m) <- c("cd62lhi","cd62lneg","B","M","E")
    return(m)
}



cellcountNormalisation <- function(mouse, mouse.mat){

    mat.cellnorm <- mouse.mat
    
    mat.cellnorm[,1] <- mouse.mat[,1] * cell_counts[cell_counts$Mouse == mouse  & cell_counts$Gate == 'cd62l-high',6]
    mat.cellnorm[,2] <- mouse.mat[,2] * cell_counts[cell_counts$Mouse == mouse & cell_counts$Gate == 'cd62l-neg',6]
    mat.cellnorm[,3] <- mouse.mat[,3] * cell_counts[cell_counts$Mouse == mouse  & cell_counts$Gate == 'B',6]
    mat.cellnorm[,4] <- mouse.mat[,4] * cell_counts[cell_counts$Mouse == mouse  & cell_counts$Gate == 'M',6]
    mat.cellnorm[,5] <- mouse.mat[,5] * cell_counts[cell_counts$Mouse == mouse  & cell_counts$Gate == 'E',6]
    
    colnames(mat.cellnorm) <- c("cd62lhi","cd62lneg","B","M","E")
    return(mat.cellnorm)

}


makeMetaMouse <- function(m1,m2,m3,m4){

  
  #rename columns so they are consistent across datasets
  colnames(m1) <- c("cd62lhi","cd62lneg","B","M","E")
  colnames(m2) <- c("cd62lhi","cd62lneg","B","M","E")
  colnames(m3) <- c("cd62lhi","cd62lneg","B","M","E")
  colnames(m4) <- c("cd62lhi","cd62lneg","B","M","E")
  
  #now make an integrated sample with all mice
  supermouse <- rbind(m1, m2,m3,m4)
  
  #clones must be in either the pos or the neg
  supermouse <- supermouse[rowSums(supermouse) > 0,]
  
  return(supermouse)
}

# set a new parameter bias stating whether its a cd62lpos or cd62l neg clone. Cells with non zero expression in both samples are considered unbiased
setCloneBias <- function(supermouse,threshold1, threshold2){
  supermouse$bias <- rep('None',nrow(supermouse))
  supermouse$bias[which(supermouse$cd62lhi <= threshold1 & supermouse$cd62lneg > threshold2)] <- "cd62lneg"
  supermouse$bias[which(supermouse$cd62lneg <= threshold1 & supermouse$cd62lhi > threshold2)] <- "cd62lhi"
  return(supermouse)
}



#perform row level normalisation to compute bias
performRowNormalisation <- function(matrix){
  for(i in 1:nrow(matrix)){
      if(rowSums(matrix[i,c("B","M","E")]) > 0){
              matrix[i,c("B","M","E")] <-   
                matrix[i,c("B","M","E")]/sum(matrix[i,c("B","M","E")])
       }
     }
  return(matrix)
}


generateCumulativeProbabilityNorm <- function(barcodematrix,sample){
  
  #renormalise the clones to 1 for the myeloid lineage
  barcodematrix$data.normalised <- barcodematrix[,sample] / sum(barcodematrix[,sample])

  #rank each clone in terms of normalised abundance
  data.ranked <- barcodematrix[order(barcodematrix$data.normalised),]
  
  output <- data.frame(cbind((data.ranked[,sample]),
                             cumsum(data.ranked$data.normalised)))
  colnames(output) <- c("cell_counts","cumsum")
  rownames(output) <- rownames(data.ranked)
  return(output)
  
}


plotDiversity <- function(df){
  barplot(c(nrow(df[df$cd62lneg > 0,]),
  nrow(df[df$cd62lhi > 0,])), main = "barcode diversity in the myeloid lineage",ylab = "number of barcodes in M",names = c("cd62Lneg","cd62Lhigh"))
}


plotClassifier <- function(mouse){
qplot(asinh(mouse$cd62lneg) ,asinh(mouse$cd62lhi),col = mouse$bias) + geom_abline(intercept = 0,slope = 1, col = 'grey30') + scale_color_manual(values=c("blue", "red","grey60"))   +  geom_point(size=3) + theme(panel.background = element_blank())
}


cloneSizeDistribution <- function(df.cellnorm,
                                  cd62lneg.barcodes,cd62lhi.barcodes,lin){

  
  df.cellnorm.subset <- df.cellnorm[c(cd62lneg.barcodes,cd62lhi.barcodes),]
  df.cellnorm.subset$bias <- "None"
  df.cellnorm.subset[cd62lneg.barcodes,]$bias <- "neg"
  df.cellnorm.subset[cd62lhi.barcodes,]$bias <- "hi"
  df.cellnorm.subset$bias <- factor(df.cellnorm.subset$bias, levels = c("neg","hi"))
  
  p <- ggplot(df.cellnorm.subset,aes_string(x = 'bias', y = lin), color = c("black","blue")) + geom_boxplot(lwd = 1.3) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + geom_jitter(dotsize = 0.5)
    
   return(p)

}


lineageBias <- function(supermouse,lin){

  supermouse.rownorm<- performRowNormalisation(supermouse)
  
  
  filtered <- supermouse.rownorm[supermouse.rownorm$bias == "cd62lneg" | supermouse.rownorm$bias == "cd62lhi",c("E","M","B","bias")]
  
  filtered$bias <- factor(filtered$bias, levels = c("cd62lneg","cd62lhi"))
  
  p <- ggplot(filtered,aes_string(x = 'bias', y = lin)) + geom_boxplot(lwd = 1.3) + geom_jitter(dotsize = 0.5)
  return(p)

}


```

### Step 2: Load the data
```{r load_data}


#load in the QC'd data matrix. 
wt.mat <- read.csv("JCW24WT_ANALYSIS_matx_norm_filt_cor_ab_trueFalseRU.csv")

#perform column-level normalisation to each sample
wt.mat.norm <- columnNormalisation(wt.mat[,-1],wt.mat$tag)

#remove barcodes that are not expressed in any of the samples
wt.mat.norm.filt <- removeUninformativeClones(wt.mat.norm)

write.csv(wt.mat.norm.filt,"/Users/jasoncosgrove/Desktop/S5.csv")
```


### Step 3: Make a count matrix for each mouse

```{r count_matrix_each_mouse}

#obtain the matrices for each mouse
wt.m1<- getMouseMatrix(wt.mat.norm.filt,"M1.WT")
wt.m2<- getMouseMatrix(wt.mat.norm.filt,"M2.WT")
wt.m3<- getMouseMatrix(wt.mat.norm.filt,"M3.WT")
wt.m4<- getMouseMatrix(wt.mat.norm.filt,"M4.WT")

#normalise the values to cellcounts
m1.wt.cellnorm <- cellcountNormalisation("M1-WT", wt.m1)
m2.wt.cellnorm <- cellcountNormalisation("M2-WT", wt.m2)
m3.wt.cellnorm <- cellcountNormalisation("M3-WT", wt.m3)
m4.wt.cellnorm <- cellcountNormalisation("M4-WT", wt.m4)

# merge the different mice into a single sample
wt.meta <- makeMetaMouse(wt.m1, wt.m2, wt.m3, wt.m4 )

tiff("/Users/jasoncosgrove/Desktop/barcode_diversity.tiff",
     width =6, height = 4, units = 'in', res = 300)
barplot(c(nrow(wt.meta[wt.meta$M > 0,]),
          nrow(wt.meta[wt.meta$B > 0,]),
          nrow(wt.meta[wt.meta$E > 0,]),
          nrow(wt.meta[wt.meta$cd62lhi > 0,]),
          nrow(wt.meta[wt.meta$cd62lneg > 0,])),
        names = c("M","B","E","cd62lhi","cd62lneg"),ylab = "number of barcodes")

dev.off()


tiff("/Users/jasoncosgrove/Desktop/barcode_diversity.tiff",
     width =8, height = 6, units = 'in', res = 300)
barplot(c(nrow(wt.meta[wt.meta$M > 0 & wt.meta$cd62lneg > 0,]),
          nrow(wt.meta[wt.meta$B > 0 & wt.meta$cd62lneg > 0,]),
          nrow(wt.meta[wt.meta$E > 0 & wt.meta$cd62lneg > 0,]),
          nrow(wt.meta[wt.meta$M > 0 & wt.meta$cd62lhi > 0,]),
          nrow(wt.meta[wt.meta$B > 0 & wt.meta$cd62lhi > 0,]),
          nrow(wt.meta[wt.meta$E > 0 & wt.meta$cd62lhi > 0,])),
        names = c("cd62lneg-M","cd62lneg-B","cd62lneg-E","cd62lhi-M","cd62lhi-B","cd62lhi-E"),ylab = "number of barcodes found across sample pairs",ylim = c(0,500))
dev.off()


# merge the different mice into a single sample
wt.meta.cellnorm <- makeMetaMouse(m1.wt.cellnorm,m2.wt.cellnorm ,m3.wt.cellnorm ,m4.wt.cellnorm )

```



### Step 3: Subset on barcodes that are only in CD62Lneg and CD62Lhi

In this section we set a read threshold for barcode classification. This allows us to see which barcodes are found in both CD62Lneg and CD62Lhi and which is found in only one of these classes. 

If we dont set a strict threshold we can get much higher number of barcodes, however a large proportion of the barcodes are very lowly expressed and so we have much less confidence that the barcodes are truly restricted to one of the CD62L subsets as sampling biases play a much larger role in lowly abundant clones. 

```{r cd62l_analysis, echo=FALSE}
tiff("/Users/jasoncosgrove/Desktop/read_thresholdhi.tiff",
     width =5, height = 5, units = 'in', res = 300)
wt.cumsum <- generateCumulativeProbabilityNorm(wt.meta,"cd62lhi")
plot(1:nrow(wt.cumsum)/nrow(wt.cumsum),wt.cumsum$cumsum,col = "blue",lwd = 1,xlab = "% of all clones", ylab = "cumulative cellular output",cex  = 0.5)
abline(h = 0.005, lwd = 2, col = "red")
dev.off()


tiff("/Users/jasoncosgrove/Desktop/read_thresholdneg.tiff",
     width =5, height = 5, units = 'in', res = 300)
wt.cumsum <- generateCumulativeProbabilityNorm(wt.meta,"cd62lneg")
plot(1:nrow(wt.cumsum)/nrow(wt.cumsum),wt.cumsum$cumsum,col = "blue",lwd = 1,xlab = "% of all clones", ylab = "cumulative cellular output",cex  = 0.5)
abline(h = 0.005, lwd = 2, col = "red")
dev.off()

#we have a lot of low abundance barcodes in progenitors
plot(wt.meta$cd62lhi)
abline(h = 0.005, lwd = 2, col = "red")
plot(wt.meta$cd62lneg)
abline(h = 0.005, lwd = 2, col = "red")

#keep only the barcodes which are well expressed in progneitors
mpp.only <- wt.meta[wt.meta$cd62lhi > 0.005 | wt.meta$cd62lneg > 0.005,]
plot(mpp.only$cd62lhi)
plot(mpp.only$cd62lneg)


setCloneBias2 <- function(supermouse,threshold1 = 0.9){
  
  
  b <- supermouse$cd62lhi/ (supermouse$cd62lhi + supermouse$cd62lneg)
  supermouse$bias <- rep('None',nrow(supermouse))
  supermouse$bias[which(b >= threshold1)] <- "cd62lhi"
  supermouse$bias[which(b <= (1 - threshold1))] <- "cd62lneg"
  return(supermouse)
}

supermouse <- setCloneBias2(mpp.only,threshold1= 0.95)
table(supermouse$bias)


nrow(mpp.only[mpp.only$cd62lhi > 0 & mpp.only$cd62lneg > 0,]) /
  nrow(mpp.only[mpp.only$cd62lhi > 0 | mpp.only$cd62lneg > 0,])

tiff("/Users/jasoncosgrove/Desktop/classifier.tiff",
     width =5, height = 4, units = 'in', res = 300)
ggplot(supermouse,aes(x = asinh(cd62lneg),y =asinh(cd62lhi), color = bias)) + geom_point(size=2) + scale_color_manual(values=c('blue','grey30', 'grey80')) +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
dev.off()



```


### Step 4: diversity, clone size and bias analysis
```{r visualise_bias_barcodes, echo=FALSE}


mean(c( nrow(wt.m1[wt.m1$cd62lhi > 0 & wt.m1$cd62lneg >0,])/
nrow(wt.m1[wt.m1$cd62lhi > 0 | wt.m1$cd62lneg >0,]),

nrow(wt.m4[wt.m2$cd62lhi > 0 & wt.m2$cd62lneg >0,])/
nrow(wt.m4[wt.m2$cd62lhi > 0 | wt.m2$cd62lneg >0,]),

nrow(wt.m4[wt.m3$cd62lhi > 0 & wt.m3$cd62lneg >0,])/
nrow(wt.m4[wt.m3$cd62lhi > 0 | wt.m3$cd62lneg >0,]),

nrow(wt.m4[wt.m4$cd62lhi > 0 & wt.m4$cd62lneg >0,])/
nrow(wt.m4[wt.m4$cd62lhi > 0 | wt.m4$cd62lneg >0,])))

sd(c( nrow(wt.m1[wt.m1$cd62lhi > 0 & wt.m1$cd62lneg >0,])/
nrow(wt.m1[wt.m1$cd62lhi > 0 | wt.m1$cd62lneg >0,]),

nrow(wt.m4[wt.m2$cd62lhi > 0 & wt.m2$cd62lneg >0,])/
nrow(wt.m4[wt.m2$cd62lhi > 0 | wt.m2$cd62lneg >0,]),

nrow(wt.m4[wt.m3$cd62lhi > 0 & wt.m3$cd62lneg >0,])/
nrow(wt.m4[wt.m3$cd62lhi > 0 | wt.m3$cd62lneg >0,]),

nrow(wt.m4[wt.m4$cd62lhi > 0 & wt.m4$cd62lneg >0,])/
nrow(wt.m4[wt.m4$cd62lhi > 0 | wt.m4$cd62lneg >0,])))


prog1 <- "cd62lhi"
prog2 <- "cd62lneg"
lin <- "M"

df <- wt.m1
df.filter <- df[df[,prog1] > 0.005 |df[,prog2] > 0.005 ,]
print(c(nrow(df.filter[df.filter[,prog] > 0 & df.filter[,lin] >0,])))

df <- wt.m2
df.filter <- df[df[,prog1] > 0.005 |df[,prog2] > 0.005 ,]
print(c(nrow(df.filter[df.filter[,prog] > 0 & df.filter[,lin] >0,])))

df <- wt.m3
df.filter <- df[df[,prog1] > 0.005 |df[,prog2] > 0.005 ,]
print(c(nrow(df.filter[df.filter[,prog] > 0 & df.filter[,lin] >0,])))

df <- wt.m4
df.filter <- df[df[,prog1] > 0.005 |df[,prog2] > 0.005 ,]
print(c(nrow(df.filter[df.filter[,prog] > 0 & df.filter[,lin] >0,])))



#subset the different clones based on their bias state
supermouse.hi<- supermouse[supermouse$bias == "cd62lhi",]
supermouse.neg <- supermouse[supermouse$bias == "cd62lneg",]
supermouse.not_shared <- supermouse[supermouse$bias != "None",]

cd62lhi.barcodes <- rownames(supermouse.hi)
cd62lneg.barcodes <- rownames(supermouse.neg)

#heatmap visualisation of the biased clones

tiff("/Users/jasoncosgrove/Desktop/heatmap.tiff",
     width =3, height = 4, units = 'in', res = 300)
heatmap(asinh(as.matrix(supermouse.not_shared[,c(1:5)] )),cexRow = 0.1,cexCol = 0.8, col=colorRampPalette(c("black","green","red"),bias = 2)(300),scale = "none",distfun = function(x) dist(x,method = 'euclidean'))
dev.off()

#analysis 1 barcode diversity in M

m.only <- supermouse[supermouse$M > 0,]

plotDiversity(m.only)





m.cs <- cloneSizeDistribution(wt.meta.cellnorm,cd62lneg.barcodes,cd62lhi.barcodes,"M")
e.cs <- cloneSizeDistribution(wt.meta.cellnorm,cd62lneg.barcodes,cd62lhi.barcodes,"E")
b.cs <- cloneSizeDistribution(wt.meta.cellnorm,cd62lneg.barcodes,cd62lhi.barcodes,"B")


write.csv(m.cs$data,"/Users/jasoncosgrove/Desktop/m_cs.csv")
write.csv(e.cs$data,"/Users/jasoncosgrove/Desktop/e_cs.csv")
write.csv(b.cs$data,"/Users/jasoncosgrove/Desktop/b_cs.csv")
#analysis 3 barcode lineage bias

#so previosuly we normalise by column which gives us the fractional representation per lineage, a normalisation that can adjust for the different sizes of the mature compartments. Now we normalise within each clone (row normalisation). SO here you are asking per clone if i contribute to 10% of all E clones, 10% of all B clones and 20% M, then you would have a bias score of 0.25,0.25 and 0.5 for E,B, and M. 

m.lb <- lineageBias(supermouse,"M")
e.lb <- lineageBias(supermouse,"E")
b.lb <- lineageBias(supermouse,"B")


write.csv(m.lb$data,"/Users/jasoncosgrove/Desktop/m_lb.csv")
write.csv(e.lb$data,"/Users/jasoncosgrove/Desktop/e_lb.csv")
write.csv(b.lb$data,"/Users/jasoncosgrove/Desktop/b_lb.csv")

```




### Step 5: lets make the threshold stricter to make sure our results are robust
```{r visualise_bias_barcodes2, echo=FALSE}

#use a strict threshold
supermouse <- setCloneBias2(mpp.only,threshold1 = 0.99, threshold2 = 0.01)

#visualise the distribution of the biased clones
table(supermouse$bias)
plotClassifier(supermouse)


#subset the different clones based on their bias state
supermouse.hi<- supermouse[supermouse$bias == "cd62lhi",]
supermouse.neg <- supermouse[supermouse$bias == "cd62lneg",]
supermouse.not_shared <- supermouse[supermouse$bias != "None",]

cd62lhi.barcodes <- rownames(supermouse.hi)
cd62lneg.barcodes <- rownames(supermouse.neg)

#heatmap visualisation of the biased clones
heatmap(log(as.matrix(supermouse.not_shared[,c(1,2,4)] + 1)) ,col=colorRampPalette(c("black","green","red"),bias = 2)(300),scale = "none",distfun = function(x) dist(x,method = 'euclidean'))


#analysis 1 barcode diversity in M

m.only <- supermouse[supermouse$M > 0,]

plotDiversity(m.only)

#analysis 2 clone size distributions, normalised to cell counts

cloneSizeDistribution(wt.meta.cellnorm,cd62lneg.barcodes,cd62lhi.barcodes)



#analysis 3 barcode lineage bias

#so previosuly we normalise by column which gives us the fractional representation per lineage, a normalisation that can adjust for the different sizes of the mature compartments. Now we normalise within each clone (row normalisation). SO here you are asking per clone if i contribute to 10% of all E clones, 10% of all B clones and 20% M, then you would have a bias score of 0.25,0.25 and 0.5 for E,B, and M. 

lineageBias(supermouse)


```



### Step 4: Focus on myeloid producing barcodes

In the G6PD analysis we set a cumulative read threshold to assign B-cell producing barcodes - lets try a similar approach here to see if our results are robust

```{r threshold_analysis, echo=FALSE}


# plot the cumulative read distribution of barcodes in Myeloid lineage
wt.cumsum <- generateCumulativeProbabilityNorm(wt.meta,"M")
plot(1:nrow(wt.cumsum)/nrow(wt.cumsum),wt.cumsum$cumsum,col = "blue",lwd = 1,xlab = "% of all clones", ylab = "cumulative cellular output")
abline(h = 0.01, col = 'red')

#keep the barcodes that contribute to 99% of all myeloid barcode read counts
# in the G6PD analysis we set this to 0.95 but here we have fewer barcodes as we have 4 less mice
myeloid.barcodes <- wt.cumsum %>% filter(cumsum > 0.05)

#keep the same read thresholds as before for consistency
supermouse <- setCloneBias(mpp.only[rownames(myeloid.barcodes),],threshold1 = 0.001, threshold2 = 0.005)
table(supermouse$bias)


#visualise the distribution of the biased clones
qplot(asinh(supermouse$cd62lneg) ,asinh(supermouse$cd62lhi),col = supermouse$bias) + geom_abline(intercept = 0,slope = 1, col = 'grey30') + scale_color_manual(values=c("blue", "red","grey60"))   +  geom_point(size=2)


#subset the different clones based on their bias state
supermouse.hi<- supermouse[supermouse$bias == "cd62lhi",]
supermouse.neg <- supermouse[supermouse$bias == "cd62lneg",]
supermouse.not_shared <- supermouse[supermouse$bias != "None",]

cd62lhi.barcodes <- rownames(supermouse.hi)
cd62lneg.barcodes <- rownames(supermouse.neg)

#heatmap visualisation of the biased clones
heatmap(log(as.matrix(supermouse.not_shared[,c(1,2,4)] + 1)) ,col=colorRampPalette(c("black","green","red"),bias = 2)(300),scale = "none",distfun = function(x) dist(x,method = 'euclidean'))


#analysis 1 barcode diversity in M

m.only <- supermouse[supermouse$M > 0,]

plotDiversity(m.only)



#analysis 2 clone size distributions, normalised to cell counts


cloneSizeDistribution(wt.meta.cellnorm,cd62lneg.barcodes,cd62lhi.barcodes)



#analysis 3 barcode lineage bias

#so previosuly we normalise by column which gives us the fractional representation per lineage, a normalisation that can adjust for the different sizes of the mature compartments. Now we normalise within each clone (row normalisation). SO here you are asking per clone if i contribute to 10% of all E clones, 10% of all B clones and 20% M, then you would have a bias score of 0.25,0.25 and 0.5 for E,B, and M. 
lineageBias(supermouse)


```




### Step 5: Assess variation of results between mice

Because our previous analyses were performed by pooling all of the data together to increase statistical power we lose information about how well our phenotype is maintained across biological replicates. To assess this factor we perform the same analyses as above but for each mouse. 

Here we see that in mouse 2

```{r individuals_m1, echo=FALSE}

wt.m1<- getMouseMatrix(wt.mat.norm.filt,"M1.WT")
wt.m2<- getMouseMatrix(wt.mat.norm.filt,"M2.WT")
wt.m3<- getMouseMatrix(wt.mat.norm.filt,"M3.WT")
wt.m4<- getMouseMatrix(wt.mat.norm.filt,"M4.WT")

#normalise the values to cellcounts
m1.wt.cellnorm <- cellcountNormalisation("M1-WT", wt.m1)
m2.wt.cellnorm <- cellcountNormalisation("M2-WT", wt.m2)
m3.wt.cellnorm <- cellcountNormalisation("M3-WT", wt.m3)
m4.wt.cellnorm <- cellcountNormalisation("M4-WT", wt.m4)

m <- wt.m1
m.cellnorm <- m1.wt.cellnorm


mpp.only <- m[m$cd62lhi > 0 | m$cd62lneg > 0,]

mouse <- setCloneBias(mpp.only,threshold1 = 0.001, threshold2 = 0.004)

#visualise the distribution of the biased clones
qplot(asinh(mouse$cd62lneg) ,asinh(mouse$cd62lhi),col = mouse$bias) + geom_abline(intercept = 0,slope = 1, col = 'grey30') + scale_color_manual(values=c("blue", "red","grey60"))   +  geom_point(size=2)


#subset the different clones based on their bias state
mouse.hi<- mouse[mouse$bias == "cd62lhi",]
mouse.neg <- mouse[mouse$bias == "cd62lneg",]
mouse.not_shared <- mouse[mouse$bias != "None",]

cd62lhi.barcodes <- rownames(mouse.hi)
cd62lneg.barcodes <- rownames(mouse.neg)

#heatmap visualisation of the biased clones
heatmap(log(as.matrix(mouse.not_shared[,c(1,2,4)] + 1)) ,col=colorRampPalette(c("black","green","red"),bias = 2)(300),scale = "none",distfun = function(x) dist(x,method = 'euclidean'))


#analysis 1 barcode diversity in M

m.only <- mouse[mouse$M > 0,]

plotDiversity(m.only)

#analysis 2 clone size distributions, normalised to cell counts

cloneSizeDistribution(m.cellnorm,cd62lneg.barcodes,cd62lhi.barcodes)

#analysis 3 barcode lineage bias

lineageBias(mouse)




```




```{r individuals_m2, echo=FALSE}

m <- wt.m2
m.cellnorm <- m2.wt.cellnorm


mpp.only <- m[m$cd62lhi > 0 | m$cd62lneg > 0,]

mouse <- setCloneBias(mpp.only,threshold1 = 0.001, threshold2 = 0.004)

#visualise the distribution of the biased clones
qplot(asinh(mouse$cd62lneg) ,asinh(mouse$cd62lhi),col = mouse$bias) + geom_abline(intercept = 0,slope = 1, col = 'grey30') + scale_color_manual(values=c("blue", "red","grey60"))   +  geom_point(size=2)


#subset the different clones based on their bias state
mouse.hi<- mouse[mouse$bias == "cd62lhi",]
mouse.neg <- mouse[mouse$bias == "cd62lneg",]
mouse.not_shared <- mouse[mouse$bias != "None",]

cd62lhi.barcodes <- rownames(mouse.hi)
cd62lneg.barcodes <- rownames(mouse.neg)

#heatmap visualisation of the biased clones
heatmap(log(as.matrix(mouse.not_shared[,c(1,2,4)] + 1)) ,col=colorRampPalette(c("black","green","red"),bias = 2)(300),scale = "none",distfun = function(x) dist(x,method = 'euclidean'))


#analysis 1 barcode diversity in M

m.only <- mouse[mouse$M > 0,]

plotDiversity(m.only)

#analysis 2 clone size distributions, normalised to cell counts

cloneSizeDistribution(m.cellnorm,cd62lneg.barcodes,cd62lhi.barcodes)

#analysis 3 barcode lineage bias
lineageBias(mouse)




```


```{r individuals_m3, echo=FALSE}

m <- wt.m3
m.cellnorm <- m3.wt.cellnorm


mpp.only <- m[m$cd62lhi > 0 | m$cd62lneg > 0,]

mouse <- setCloneBias(mpp.only,threshold1 = 0.001, threshold2 = 0.004)

#visualise the distribution of the biased clones
qplot(asinh(mouse$cd62lneg) ,asinh(mouse$cd62lhi),col = mouse$bias) + geom_abline(intercept = 0,slope = 1, col = 'grey30') + scale_color_manual(values=c("blue", "red","grey60"))   +  geom_point(size=2)


#subset the different clones based on their bias state
mouse.hi<- mouse[mouse$bias == "cd62lhi",]
mouse.neg <- mouse[mouse$bias == "cd62lneg",]
mouse.not_shared <- mouse[mouse$bias != "None",]

cd62lhi.barcodes <- rownames(mouse.hi)
cd62lneg.barcodes <- rownames(mouse.neg)

#heatmap visualisation of the biased clones
heatmap(log(as.matrix(mouse.not_shared[,c(1,2,4)] + 1)) ,col=colorRampPalette(c("black","green","red"),bias = 2)(300),scale = "none",distfun = function(x) dist(x,method = 'euclidean'))


#analysis 1 barcode diversity in M

m.only <- mouse[mouse$M > 0,]

plotDiversity(m.only)

#analysis 2 clone size distributions, normalised to cell counts

cloneSizeDistribution(m.cellnorm,cd62lneg.barcodes,cd62lhi.barcodes)



lineageBias(mouse)



```


```{r individuals_m4, echo=FALSE}

m <- wt.m4
m.cellnorm <- m4.wt.cellnorm


mpp.only <- m[m$cd62lhi > 0 | m$cd62lneg > 0,]

mouse <- setCloneBias(mpp.only,threshold1 = 0.001, threshold2 = 0.004)

#visualise the distribution of the biased clones
qplot(asinh(mouse$cd62lneg) ,asinh(mouse$cd62lhi),col = mouse$bias) + geom_abline(intercept = 0,slope = 1, col = 'grey30') + scale_color_manual(values=c("blue", "red","grey60"))   +  geom_point(size=2)


#subset the different clones based on their bias state
mouse.hi<- mouse[mouse$bias == "cd62lhi",]
mouse.neg <- mouse[mouse$bias == "cd62lneg",]
mouse.not_shared <- mouse[mouse$bias != "None",]

cd62lhi.barcodes <- rownames(mouse.hi)
cd62lneg.barcodes <- rownames(mouse.neg)

#heatmap visualisation of the biased clones
heatmap(log(as.matrix(mouse.not_shared[,c(1,2,4)] + 1)) ,col=colorRampPalette(c("black","green","red"),bias = 2)(300),scale = "none",distfun = function(x) dist(x,method = 'euclidean'))


#analysis 1 barcode diversity in M

m.only <- mouse[mouse$M > 0,]

plotDiversity(m.only)

#analysis 2 clone size distributions, normalised to cell counts

cloneSizeDistribution(m.cellnorm,cd62lneg.barcodes,cd62lhi.barcodes)


lineageBias(mouse)



```